Supplement to Effect of COVID-19 burden on changes in HPV vaccination coverage among females between 2019 and 2022

### 2025-04-16 code/processing-code/processing-code.qmd
Part 3 Commit 1
Part 5 Commit 1

#### HPV coverage and vaccination program data
There are two data files to be imported for the HPV vaccination data, one file contains yearly coverage estimates and the other contains information about the HPV vaccination program.

##### HPV vaccination coverage
A total of 196 countries are included in the data set. Coverage estimates are available from 2010 to 2023, but most countries are missing on some years of data. Data manipulation will be needed to restrict to the years of interest and convert data form long to wide.

@@

##### HPV vaccination program characteristics
A total of 196 countries are included in the data set. About one third of countries are missing year of introduction, and a few are missing country income. Other characteristics (e.g., delivery strategy) appear complete.
```{r}
#Importing HPV vaccination program characteristics
hpv_program_raw <- read.csv(here("data","raw-data","who-dashboard-1dosecoverage-1-27-25.csv"), header = TRUE)

@@

I am incrementally building the combined data set. As a first step, I am combining the HPV vaccination coverage estimates with the HPV vaccination program information and performing some basic data checks. Other combinations will be performed in the _Combining all datasets_ section below.
```{r}

@@

#### United Nations data 
There are 238 countries with yearly data from 1950 to 2023. Data appear mostly complete (ISO2 Alpha-code is missing for some rows but this is missing for all World and Region specific rows since this is a code to indicate country). Data will be restricted to the two years of interest during the next part of the project.
```{r}

@@

#### COVID19 data 
This data source includes weekly COVD case count data on 239 countries. The data set includes 440640 rows, but about half have missing data on new case counts. Completeness of data will have to be reviewed once the data are restricted to countries with vaccination coverage estimates in the years of interest.
```{r}
#Importing HPV vaccination program characteristics
covid_raw <- read.csv(here("data","raw-data","WHO-COVID-19-global-daily-data.csv"), header = TRUE)

@@

## Combining all datasets
Above, HPV vaccination coverage and vaccine program characteristic data were combined. Here, I'm combining that data with the United Nations demographic characteristics and COVID burden.

@@

>>>>

#### HPV coverage and vaccination program data

There are two data files to be imported for the HPV vaccination data, one file contains yearly coverage estimates and the other contains information about the HPV vaccination program.

##### HPV vaccination coverage

A total of 196 countries are included in the data set. Coverage estimates are available from 2010 to 2023, but most countries are missing on some years of data. Data manipulation will be needed to restrict to the years of interest and convert data form long to wide.

@@

##### HPV vaccination program characteristics

A total of 196 countries are included in the data set. About one third of countries are missing year of introduction, and a few are missing country income. Other characteristics (e.g., delivery strategy) appear complete.

```{r}
#Importing HPV vaccination program characteristics
hpv_program_raw <- read.csv(here("data","raw-data","who-dashboard-1dosecoverage-1-27-25.csv"), header = TRUE)

@@

I am incrementally building the combined data set. As a first step, I am combining the HPV vaccination coverage estimates with the HPV vaccination program information and performing some basic data checks. Other combinations will be performed in the *Combining all datasets* section below.

```{r}

@@

#### United Nations data

There are 238 countries with yearly data from 1950 to 2023. Data appear mostly complete (ISO2 Alpha-code is missing for some rows but this is missing for all World and Region specific rows since this is a code to indicate country). Data will be restricted to the two years of interest during the next part of the project.

```{r}

@@

#### COVID19 data

This data source includes weekly COVD case count data on 239 countries. The data set includes 440640 rows, but about half have missing data on new case counts. Completeness of data will have to be reviewed once the data are restricted to countries with vaccination coverage estimates in the years of interest.

```{r}
#Importing HPV vaccination program characteristics
covid_raw <- read.csv(here("data","raw-data","WHO-COVID-19-global-daily-data.csv"), header = TRUE)

@@

## Combining all datasets

Above, HPV vaccination coverage and vaccine program characteristic data were combined. Here, I'm combining that data with the United Nations demographic characteristics and COVID burden.

@@

project_data <- project_data %>%
  mutate(delivery_strat = if_else(HPV_PRIM_DELIV_STRATEGY == "Varies by region/province", "Mixed", HPV_PRIM_DELIV_STRATEGY))

### 2025-04-16 products/manuscript/Manuscript.qmd
Part 5 Commit 1

Human papillomavirus vaccines have been available since 2006, and as of 2019, 108 countries had introduced HPV vaccine into their national immunization programs. In 2020, immunization services globally experienced disruptions due to the COVID-19 pandemic. My objective was to assess the association between country-level COVID burden and changes in HPV vaccination coverage between 2019 and 2022.

@@

Data on annual coverage, cumulative number of COVID cases between Jan. 1, 2020, and Dec. 31, 2022, country income level, vaccination program characteristics, and sociodemographic information were publicly available from the World Health Organization and United Nations. Multivariable linear regression models, modeling the difference in 2022 and 2019 HPV vaccination coverage, were performed with number of cumulative COVID cases as the primary predictor; additional country vaccination program and sociodemographic characteristics were considered for adjustment. Analyses were restricted to complete cases (N = 98). LASSO regression and regression tree models were also considered. Models were compared using R^2^ values.

@@

There was a weak, positive, linear association between the difference in HPV vaccination coverage and number of cumulative COVID cases (𝛽(se) = 0.000769 (0.00036), p-value = 0.034). Variability in the coverage difference was highest for countries with the lowest COVID burden and decreased as COVID burden increased. After adjustment for all vaccination program and sociodemographic characteristics, coverage difference and COVID burden were not associated (beta = -0.000065 (0.00064), p-value = 0.919). COVID burden was an important predictor of difference in coverage among countries with a median resident age of \<26 years.

@@

In December 2019, cases of a severe respiratory disease increased rapidly in China. We now know this was the start of the COVID-19 pandemic.[@cucinotta2020] Globally, this pandemic disrupted many medical services, including immunization programs.[@casey2024] Several studies have demonstrated the country-specific impact of the COVID-19 pandemic on HPV vaccination and HPV-related disease.[@ivanus2021; @gountas2023; @silva2022; @rao2022; @daniels2021] Additionally, a global assessment of the effect of the pandemic on HPV vaccination coverage has been performed.[@casey2024] However, these analyses focused on pre- and post-pandemic comparisons rather than considering the magnitude of COVID burden in the country. To address this gap, my objective was to assess the association between country-level COVID burden and changes in HPV vaccination coverage between 2019 and 2022. I hypothesized that decreases in HPV vaccination coverage between 2022 and 2019 were greater among countries with high burden of COVID-19 in 2020-2022. A secondary objective of this work was to develop a regression tree to assist with classifying countries by their change in HPV vaccination coverage using other country-level HPV vaccine program and sociodemographic characteristics. The purpose of this regression tree is to identify key sociodemographic indicators that may be used by policy makers for pandemic preparedness efforts to prevent reductions in HPV coverage if a new epidemic or pandemic were to affect a country.

@@

Data on HPV vaccination coverage and vaccination program details are available at the country level through the World Health Organization's (WHO) [HPV Dashboard](https://www.who.int/teams/immunization-vaccines-and-biologicals/diseases/human-papillomavirus-vaccines-(HPV)/hpv-clearing-house/hpv-dashboard).[@HPVdashboard] Briefly, coverage data reflect administrative and official HPV vaccination coverage reported annually through the WHO/UNICEF Joint Reporting Form on Immunization. The HPV Dashboard includes country income level (low, lower middle, upper middle, high), whether there was a national HPV vaccination recommendation, year of vaccine introduction, primary delivery strategy (school-based, facility based, varies by region, mixed), number of doses in routine schedule, whether the program is gender neutral (i.e., targets males and females), WHO region, and year-specific coverage from 2010 to 2023.

@@

COVID burden between January 4, 2020, and December 31, 2022, was defined as the cumulative number of COVID cases per 100,000 persons. Additionally, cumulative number of COVID deaths per 100,000 was also calculated.

@@

Univariate analyses of the outcome and each country-level characteristic were performed. A histogram of the difference in coverage was assessed to determine the approximate distribution of the outcome. The mean (standard deviation (sd)) and median (interquartile range (IQR)) were calculated for each continuous country characteristic (e.g., number of cumulative COVID cases). For categorical variables, the distribution (n, %) of each level was calculated. Bivariate associations between each characteristic and the change in HPV vaccination coverage between 2022 and 2019 were visually assessed using scatterplots (with a LOESS line of best fit for visualization purposes only) for continuous variables and violin plots for categorical variables.

@@

Multivariable linear regression was performed using the lm function to assess the association between the difference in HPV vaccination coverage and cumulative number of COVID cases, a measure of COVID burden. An unadjusted model and fully adjusted model (adjusted for all country level demographic and vaccination program characteristics) were performed. Additionally, models adjusted for each country level characteristic individually were performed. For each model, the effect of COVID case count on the difference in HPV vaccination coverage between 2019-2022 (beta, standard error, p-value from t-test with a null hypothesis of beta = 0) and model performance metrics (R-squared, AIC, BIC) were extracted and compared. Selection of a final model was based on subject matter expertise, which model had the highest R-squared, and which model had the lowest AIC and/or BIC. Model parameters were presented for the final selected model. Because there are many potential predictors, LASSO regression was also performed to reduce the risk of overfitting. LASSO regression was fit for a grid of penalty values from 1E-5 to 1E2 on a log scale using the glmnet engine and tidymodels framework. The highest tuning value without a reduction in R^2^ was selected as the final tuning value. Beta estimates and R^2^ values from the final multivariable linear regression and LASSO regression models were compared. Diagnostic plots (e.g., predicted vs observed) for both models were visually inspected to assess model fit.

@@

```{r, include = F, echo = F}
table1_display=readRDS(here("results","tables","table1_tib.rds"))
table1_tib <- as_tibble(table1_display) %>% rename(character = `Characteristic`, stat = `n/N (%), mean (sd), or median (IQR), N = 100`)
```

@@

Between January 1, 2020, and December 31, 2022, the mean (sd) number of cases per 100,000 persons was `r table1_tib %>%  filter(row_number() == which(character == "Cumulative_case_rate") + 1) %>% pull(stat)`. The median (IQR) was lower at `r table1_tib %>% filter(row_number() == which(character == "Cumulative_case_rate") + 2) %>% pull(stat)`.(@tbl-Table1_coverage_covid) The mean 3-year COVID death rate was `r table1_tib %>% filter(row_number() == which(character == "Cumulative_death_rate") + 1) %>% pull(stat)` deaths per 100,000 persons; the median was similar at `r table1_tib %>% filter(row_number() == which(character == "Cumulative_death_rate") + 2) %>% pull(stat)`.

@@

```{r}
#| label: fig-Figure2
#| fig-cap: "Scatterplot of difference in HPV vaccination coverage between 2019 and 2022 and cumulative COVID cases per 100,000"
#| echo: FALSE
knitr::include_graphics(here("results","figures","scatter_case_covdiff.png"))

@@

There were some differences in the distribution of coverage difference by vaccine program characteristics (@fig-Figure5). Countries with a facility-based delivery strategy, gender neutral vaccination program, and 2-dose recommendation had distributions of difference in coverage with a smaller portion of countries in the distribution region indicating a decrease in coverage (i.e., most of the data was centered around zero or was greater than 0). There was not a strong association with difference in coverage and year of implementation.

```{r}
#| label: fig-Figure5
#| fig-cap: "Associations between vaccine program characteristics and difference in HPV vaccination coverage between 2022 and 2019 and cumulative COVID deaths per 100,000"
#| echo: FALSE
knitr::include_graphics(here("results","figures","bivar_vac.png"))
```

@@

For the unadjusted and all adjusted multivariable linear regression models, COVID burden beta estimates, standard errors, p-values for the effect of cumulative COVID cases on the difference in coverage, and performance metrics (R^2^, AIC, BIC) are presented in @tbl-modelcomp. Each performance metric has been shaded using a sliding color scale; a light color indicates worse performance, darker color indicates better performance, and the "best" performing model is indicated using bold text. The fully adjusted model (adjusted for all country-level demographic and vaccination program characteristics) had the highest R^2^. The model adjusted for median resident age had the lowest AIC and BIC. The association between change in coverage and COVID burden were similar in the fully adjusted model (𝛽=`r prelim_lm_betas %>% filter(row_number() == which(Adjustment == "Full model")) %>% pull(beta) %>% format(scientific = FALSE)`) and model only adjusted for median age (𝛽=`r prelim_lm_betas %>% filter(row_number() == which(Adjustment == "Median age")) %>% pull(beta) %>% format(scientific = FALSE)`) (@tbl-modelcomp, see supplementary material for all model parameters for median age adjusted model). Preliminary modeling will focus on the fully adjusted model. However, none of these models fit particularly well based on the low R^2^ values and diagnostic plots in @fig-lm-diagnostics. Based on the predicted vs observed plot, many observations with relatively high or low observed coverage difference values are further from the line of perfect fit. Additionally, there appears to be a slight pattern in the residuals; relatively small predicted values are more likely to have negative residuals and moderate values are more likely to have large positive values. Given that the beta is switching directions (positive/negative) based on the adjustment variables and is very close to zero (@tbl-modelcomp), there does not appear to be a strong association between COVID burden and difference in coverage regardless of adjustment.

@@

Because of the large number of coefficients in the full model, LASSO regression model was also performed for regularization. A grid of penalty values from 1E-5 to 1E2 on a log scale was considered for tuning. @Fig-Lasso_fit shows the R^2^ value for each model tuning parameter in the left panel, and the residuals from the run with each tuning parameter in the right panel. Each color represents a particular run. R^2^ is relatively stable for tuning parameters values from 1E-5 to 1E-2; therefore, a tuning value of 1E-2 was selected for the final model.

@@

Estimates from the fully adjusted model are shown in @tbl-fully_adj. There was not a significant association between cumulative number of COVID cases and the difference in HPV vaccination coverage between 2019-2022 using the linear model (𝛽(se) = `r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(estimate) %>% format(scientific = FALSE)` (`r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(std.error) %>% format(scientific = FALSE)`), *p*-value = `r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(p.value)`). After adjusting for all other characteristics, the difference in coverage was significantly higher and positive (indicating increase in coverage between 2019 and 2022) for all WHO regions, except the Eastern Mediterranean Region, compared to the African Region. Conclusions from the LASSO regression model were similar. However, some beta estimates other than COVID burden changed magnitude and/or direction (e.g., population growth rate, low income) compared to the linear regression model.

```{r}
#| label: tbl-fully_adj
#| tbl-cap: "Model parameters for fully adjusted model assessing the effect of cumulative COVID cases on the difference in HPV vaccination coverage"
#| echo: FALSE
knitr::include_graphics(here("results","tables","lmtablefull_gt.png"))

@@

Of the four models considered, the regression tree had the largest R^2^ value (`r r_sq_table_ref %>% filter(row_number() == which(Model == "Regression tree")) %>% pull(rsq_est) %>% round(3)`). This model is also potentially the most intuitive, and the utility of estimating COVID burden is apparent in this model.

>>>>

Human papillomavirus vaccines have been available since 2006, and as of 2019, 108 countries had introduced HPV vaccine into their national immunization programs. In 2020, immunization services, globally, experienced disruptions due to the COVID-19 pandemic. My objective was to assess the association between country-level COVID burden and changes in HPV vaccination coverage between 2019 and 2022.

@@

Data on annual coverage, cumulative number of COVID cases between Jan. 1, 2020, and Dec. 31, 2022, country income level, vaccination program characteristics, and sociodemographic information were publicly available from the World Health Organization and United Nations. Multivariable linear regression models, modeling the difference in 2022 and 2019 HPV vaccination coverage, were performed with number of cumulative COVID cases as the primary predictor; additional country vaccination program and sociodemographic characteristics were considered for adjustment. Analyses were restricted to complete cases (N = 98). LASSO regression and regression tree models were also considered.

@@

There was a weak, positive, linear association between the difference in HPV vaccination coverage and number of cumulative COVID cases (𝛽(se) = 0.000769 (0.00036), p-value = 0.034). Variability in the coverage difference was highest for countries with the lowest COVID burden and decreased as COVID burden increased. After adjustment for all vaccination program and sociodemographic characteristics, coverage difference and COVID burden were not associated (𝛽= -0.000065 (0.00064), p-value = 0.919). Based on the regression tree analysis, COVID burden was an important predictor of difference in coverage among countries with a median resident age of \<26 years.

@@

In December 2019, cases of a severe respiratory disease increased rapidly in China. We now know this was the start of the COVID-19 pandemic.[@cucinotta2020] Globally, this pandemic disrupted many medical services, including immunization programs.[@casey2024] Several studies have demonstrated the country-specific impact of the COVID-19 pandemic on HPV vaccination and HPV-related disease.[@ivanus2021; @gountas2023; @silva2022; @rao2022; @daniels2021] Additionally, a global assessment of the effect of the pandemic on HPV vaccination coverage has been performed.[@casey2024] However, these analyses focused on pre- and post-pandemic comparisons rather than considering the magnitude of COVID burden in the country. To address this gap, my objective was to assess the association between country-level COVID burden and changes in HPV vaccination coverage between 2019 and 2022. I hypothesized that decreases in HPV vaccination coverage between 2019 and 2022 would be greater among countries with high burden of COVID-19 in 2020-2022. A secondary objective of this work was to develop a regression tree to assist with classifying countries by their change in HPV vaccination coverage using other country-level HPV vaccine program and sociodemographic characteristics. The purpose of this regression tree is to identify key sociodemographic indicators that may be used by policy makers for pandemic preparedness efforts to prevent reductions in HPV vaccination coverage if a new epidemic or pandemic were to affect a country.

@@

Data on HPV vaccination coverage and vaccination program details are available at the country level through the World Health Organization's (WHO) [HPV Dashboard](https://www.who.int/teams/immunization-vaccines-and-biologicals/diseases/human-papillomavirus-vaccines-(HPV)/hpv-clearing-house/hpv-dashboard).[@HPVdashboard] Briefly, coverage data reflect administrative and official HPV vaccination coverage reported annually through the WHO/UNICEF Joint Reporting Form on Immunization. The HPV Dashboard includes country income level (low, lower middle, upper middle, high), whether there was a national HPV vaccination recommendation, year of vaccine introduction, primary delivery strategy (school-based, facility based, varies by region, mixed), number of doses in routine schedule (1 or 2), whether the program is gender neutral (i.e., targets males and females or females only), WHO region, and year-specific coverage from 2010 to 2023.

@@

COVID burden between January 4, 2020, and December 31, 2022, was defined as the cumulative number of COVID cases per 100,000 persons. Additionally, the cumulative number of COVID deaths per 100,000 persons was also calculated.

@@

Univariate analyses of the outcome and each country-level characteristic were performed. A histogram of the difference in coverage was assessed to determine the approximate distribution of the outcome. The mean (standard deviation (sd)) and median (interquartile range (IQR)) were calculated for each continuous country characteristic (e.g., number of cumulative COVID cases). For categorical variables, the distribution (n, %) of each level was calculated. Bivariate associations between each characteristic and the difference in HPV vaccination coverage between 2022 and 2019 were visually assessed using scatterplots (with a LOESS line of best fit for visualization purposes only) for continuous variables and violin plots for categorical variables.

@@

Multivariable linear regression was performed using the lm function to assess the association between the difference in HPV vaccination coverage and cumulative number of COVID cases, a measure of COVID burden. An unadjusted model and fully adjusted model (adjusted for all country level demographic and vaccination program characteristics) were performed. Additionally, models adjusted for each country level characteristic individually were performed. For each model, the effect of COVID case count on the difference in HPV vaccination coverage between 2022 and 2019 (beta, standard error, p-value from t-test with a null hypothesis of beta = 0) and model performance metrics (R-squared, AIC, BIC) were extracted and compared. Selection of a final model was based on subject matter expertise, which model had the highest R-squared, and which model had the lowest AIC and/or BIC. Model parameters were presented for the final selected model. Because there are many potential predictors, LASSO regression was also performed to reduce the risk of overfitting. LASSO regression was fit for a grid of penalty values from 1E-5 to 1E2 on a log scale using the glmnet engine and tidymodels framework. The highest tuning value without a reduction in R^2^ was selected as the final tuning value. Beta estimates and R^2^ values from the final multivariable linear regression and LASSO regression models were compared. Diagnostic plots (e.g., predicted vs observed) for both models were visually inspected to assess model fit.

@@

```{r, include = F, echo = F}
table1_display=readRDS(here("results","tables","table1_tib.rds"))
table1_tib <- as_tibble(table1_display) %>% rename(character = `Characteristic`, stat = `n/N (%), mean (SD), or median (IQR), N = 98`)
```

@@

Between January 1, 2020, and December 31, 2022, the mean (sd) number of COVID cases per 100,000 persons was `r table1_tib %>%  filter(row_number() == which(character == "Cumulative_case_rate") + 1) %>% pull(stat)`. The median (IQR) was lower at `r table1_tib %>% filter(row_number() == which(character == "Cumulative_case_rate") + 2) %>% pull(stat)`.(@tbl-Table1_coverage_covid) The mean 3-year COVID death rate was `r table1_tib %>% filter(row_number() == which(character == "Cumulative_death_rate") + 1) %>% pull(stat)` deaths per 100,000 persons; the median was similar at `r table1_tib %>% filter(row_number() == which(character == "Cumulative_death_rate") + 2) %>% pull(stat)`.

@@

```{r}
#| label: fig-Figure2
#| fig-cap: "Scatterplot of difference in HPV vaccination coverage between 2022 and 2019 and cumulative COVID cases per 100,000"
#| echo: FALSE
knitr::include_graphics(here("results","figures","scatter_case_covdiff.png"))

@@

There were some differences in the distribution of coverage difference by vaccine program characteristics (@fig-Figure5). Countries with a facility-based delivery strategy, gender neutral vaccination program, and 2-dose recommendation had distributions of difference in coverage with a smaller portion of countries in the distribution region indicating a decrease in coverage (i.e., most of the data was centered around zero or was greater than 0). There was not a strong association between difference in coverage and year of implementation.

```{r}
#| label: fig-Figure5
#| fig-cap: "Associations between vaccine program characteristics and difference in HPV vaccination coverage between 2022 and 2019 and vaccination program characteristics"
#| echo: FALSE
knitr::include_graphics(here("results","figures","bivar_vac.png"))
```

@@

For the unadjusted and all adjusted multivariable linear regression models, COVID burden beta estimates, standard errors, and p-values for the effect of cumulative COVID cases on the difference in coverage as well as model performance metrics (R^2^, AIC, BIC) are presented in @tbl-modelcomp. Each performance metric has been shaded using a sliding color scale; a light color indicates worse performance, darker color indicates better performance, and the "best" performing model is indicated using bold text. The fully adjusted model (adjusted for all country-level demographic and vaccination program characteristics) had the highest R^2^. The model adjusted for median resident age had the lowest AIC and BIC. The association between change in coverage and COVID burden was similar in the fully adjusted model (𝛽=`r prelim_lm_betas %>% filter(row_number() == which(Adjustment == "Full model")) %>% pull(beta) %>% format(scientific = FALSE)`) and model only adjusted for median age (𝛽=`r prelim_lm_betas %>% filter(row_number() == which(Adjustment == "Median age")) %>% pull(beta) %>% format(scientific = FALSE)`) (@tbl-modelcomp, see supplementary material for all model parameters for median age adjusted model). Multivariable linear regression modeling will focus on the fully adjusted model. However, none of these models fit particularly well based on the low R^2^ values and diagnostic plots in @fig-lm-diagnostics. Based on the predicted vs observed plot, many observations with relatively high or low observed coverage difference values are further from the line of perfect fit. Additionally, there appears to be a slight pattern in the residuals; relatively small predicted values are more likely to have negative residuals and moderate values are more likely to have large positive values. Given that the beta is switching directions (positive/negative) based on the adjustment variables and is very close to zero (@tbl-modelcomp), there does not appear to be a strong association between COVID burden and difference in coverage regardless of adjustment.

@@

Because of the large number of coefficients in the full model, LASSO regression was also performed for regularization. A grid of penalty values from 1E-5 to 1E2 on a log scale was considered for tuning. @Fig-Lasso_fit shows the R^2^ value for each model tuning parameter in the left panel, and the residuals from the run with each tuning parameter in the right panel. Each color represents a particular run. R^2^ is relatively stable for tuning parameter values from 1E-5 to 1E-2; therefore, a tuning value of 1E-2 was selected for the final model.

@@

Estimates from the fully adjusted model are shown in @tbl-fully_adj. There was not a significant association between cumulative number of COVID cases and the difference in HPV vaccination coverage between 2019-2022 using the linear model (𝛽(se) = `r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(estimate) %>% format(scientific = FALSE)` (`r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(std.error) %>% format(scientific = FALSE)`), *p*-value = `r reg_betas %>% filter(row_number() == which(term == "Cumulative_case_rate")) %>% pull(p.value)`). After adjusting for all other characteristics, the difference in coverage was significantly higher and positive (indicating an increase in coverage between 2019 and 2022) for all WHO regions, except the Eastern Mediterranean Region, compared to the African Region. Conclusions from the LASSO regression model were similar. However, some beta estimates other than COVID burden changed magnitude and/or direction (e.g., population growth rate, low income) compared to the linear regression model.

{{< pagebreak >}}
```{r}
#| label: tbl-fully_adj
#| tbl-cap: "Model parameters for fully adjusted model assessing the effect of cumulative COVID cases on the difference in HPV vaccination coverage"
#| echo: FALSE
knitr::include_graphics(here("results","tables","lmtablefull_gt.png"))

@@

Of the four models considered, the regression tree had the largest R^2^ value (`r r_sq_table_ref %>% filter(row_number() == which(Model == "Regression tree")) %>% pull(rsq_est) %>% round(3)`; @tbl-rsq-models). This model is also potentially the most intuitive, and the utility of estimating COVID burden is apparent in this model.

### 2025-04-16 products/manuscript/supplement/Supplementary-Material.qmd
Part 5 Commit 1

* The `data` folder contains all data files required for the project. There are 2 sub-folders, raw-data contains raw data files
  downloaded from their online sourcesand processed-data which contains the cleaned and edited data sets, including a final
  merged project dataset (project_data.rds). All files are saved as .rds files.

* The `products` folder contains a subfolder for the manuscript, 
  - The  `manuscript` subfolder contains a sub-folder for a supplementary material file.

* The `results` folder contains automatically/code generated output. This includes subfolders for figures and tables. 
  All figures are saved as .png files. Most tables are saved as .png and .rds files. Rds files can be used to reference specific 
  values, while the png files will be displayed in the body of the manuscript since Quarto is currently unable to knit tables to 
  Word. All content in these folders are automatically generated by code.

* There are multiple special files in the repo.
  * `readme.md`: this file contains instructions or details about the folder it
  is located in. You are reading the project-level `README.md` file right now. There is a `readme` in almost every folder.
  * `data-analysis-template.Rproj` is a file that tells RStudio that this is the main folder for a project. 
  * a few "hidden" files and folders (they start with a `.` and depending on how your OS is configured, you might not see them). Those are for R/RStudio and Git/GitHub and you can ignore them.

@@

-   HPV vaccination coverage (WHO)
    -   HPV vaccination coverage in 2022
    -   HPV vaccination coverage in 2019
-   HPV vaccination program characteristics (WHO)
    -   WHO region
    -   World Bank country income level
    -   Indicator for whether HPV vaccine is offered through a national recommendation
    -   Year of HPV vaccine program introduction
    -   Primary HPV vaccine delivery strategy (e.g., school-based)
    -   Indicator for gender neutral vaccination (i.e., female-only vaccination or females and males)
    -   Recommended number of doses (1 or 2)
-   United Nations country demographic data (UN)
    -   Total Population, as of 1 July (thousands)
    -   Population Density, as of 1 July (persons per square km)
    -   Median Age, as of 1 July (years)
    -   Population Growth Rate (percentage)
    -   Total Fertility Rate (live births per woman)
    -   Crude Death Rate (deaths per 1,000 population)
    -   Life Expectancy at Birth, both sexes (years)
    -   Net Migration Rate (per 1,000 population)
-   COVID burden through Dec. 31, 2022 (WHO)
    -   Cumulative number of COVID cases per 100,000 persons
    -   Cumulative number of COVID deaths per 100,000 persons

@@

```{r}
#| label: fig-Figure
#| fig-cap: "Scatterplot of difference in HPV vaccination coverage between 2019 and 2022 and cumulative COVID deaths per 100,000"
#| echo: FALSE
knitr::include_graphics(here("results","figures","scatter_death_covdiff.png"))
```

>>>>

* The `data` folder contains all data files required for the project. There are 2 sub-folders: raw-data contains raw data files
  downloaded from their online sources and processed-data contains the cleaned and edited data sets, including a final
  merged project data set (project_data.rds). All files are saved as .rds files.

* The `products` folder contains a subfolder for the manuscript, 
  - The  `manuscript` subfolder contains a subfolder for a supplementary material file.

* The `results` folder contains code generated output and includes subfolders for figures and tables. 
  All figures are saved as .png files. Most tables are saved as .png and .rds files. Rds files can be used to reference specific 
  values, while the png files will be displayed in the body of the manuscript since Quarto is currently unable to knit tables to 
  Word. All content in these folders are automatically generated by code.

* There are multiple special files in the repo.
  * `readme.md`: this file contains instructions and/or details about the folder it is located in. 
    The information you are reading now is covered in the project-level `README.md` file. There is a `readme` in almost every folder.
  * `data-analysis-template.Rproj` is a file that tells RStudio that this is the main folder for a project. 
  * a few "hidden" files and folders (they start with a `.` and depending on how your OS is configured, you might not see them). Those are for R/RStudio and Git/GitHub and you can ignore them.

@@

-   HPV vaccination coverage (WHO)
    -   HPV vaccination coverage in 2022
    -   HPV vaccination coverage in 2019
-   HPV vaccination program characteristics (WHO)
    -   WHO region
    -   World Bank country income level
    -   Indicator for whether HPV vaccine is offered through a national recommendation
    -   Year of HPV vaccine program introduction
    -   Primary HPV vaccine delivery strategy (school-based, clinic-based, varies)
    -   Indicator for gender neutral vaccination (female-only vaccination or females and males)
    -   Recommended number of doses (1 or 2)
-   United Nations country demographic data (UN)
    -   Total Population, as of 1 July (thousands)
    -   Population Density, as of 1 July (persons per square km)
    -   Median Age, as of 1 July (years)
    -   Population Growth Rate (percentage)
    -   Total Fertility Rate (live births per woman)
    -   Crude Death Rate (deaths per 1,000 population)
    -   Life Expectancy at Birth, both sexes (years)
    -   Net Migration Rate (per 1,000 population)
-   COVID burden through Dec. 31, 2022 (WHO)
    -   Cumulative number of COVID cases per 100,000 persons
    -   Cumulative number of COVID deaths per 100,000 persons

@@

```{r}
#| label: fig-Figure
#| fig-cap: "Scatterplot of difference in HPV vaccination coverage between 2022 and 2019 and cumulative COVID deaths per 100,000 persons"
#| echo: FALSE
knitr::include_graphics(here("results","figures","scatter_death_covdiff.png"))
```
